---
title: "Baseline analysis"
format: html
editor: visual
execute: 
  keep-md: true
---

```{r, echo=FALSE, message=FALSE}
theme_set(theme_bw())
```

## Read data

```{r}
df <- tar_read(baseline)
df <- rename(df, percentage_sharing = ..sharing)
```

## Analyse parameters

Parameters which are varied are:

-   sharing-incentive
-   mandate-sharing?
-   originator benefit

The rest is not changed at all for now.

Outcome variables are:

-   gini of resources
-   mean effort of resources
-   total sum of resources
-   \% of teams sharing

```{r}
outcomes <- c("gini..resources..of.teams", "mean..effort..of.teams",
         "sum..resources..of.teams", "percentage_sharing")

df_small <- df %>% 
  select(.run.number., .step., sharing.incentive, mandate.sharing., 
         originator.benefit, all_of(outcomes))
```

### Vary sharing incentive

Here we hold originator benefit constant at .1, and contrast sharing mandates
(sharing mandate is either )

```{r}
incentive_outcomes <- df_small %>% 
  filter(originator.benefit == .1) %>% 
  group_by(sharing.incentive, .step., mandate.sharing.) %>% 
  summarise(across(all_of(outcomes), mean))
```

```{r incentives-mandates, fig.width=8, fig.height=8}
incentive_outcomes %>% 
  pivot_longer(-c(sharing.incentive, .step., mandate.sharing.),
               names_to = "outcome") %>% 
  mutate(mandate.sharing. = ifelse(mandate.sharing., 
                                   "Sharing mandate", "No mandate")) %>% 
  ggplot(aes(.step., value, colour = as.factor(sharing.incentive))) +
  geom_line() +
  facet_grid(rows = vars(outcome),
             cols = vars(mandate.sharing.), scales = "free_y") +
  labs(colour = "Sharing incentive") +
  theme(legend.position = "top") +
  guides(colour = guide_legend(nrow = 1))
```

This is highly peculiar: If sharing is not mandated, incentives work as
anticipated: higher incentives lead to higher sharing (but overall still low
sharing, up to about 30% of groups sharing). If sharing is mandated, no
incentives lead to the highest sharing percentage.

Another issue that seems to be a coding error is observed when looking at the
sum of resources: these are very low for "No mandate", although data is being
shared. This is very unexpected (redistribution of resources should not depend
on the mandate, only on sharing behaviour).
