---
title: "Funder selectivity and inequality in data sharing"
format: 
  html:
    code-fold: true
  #pdf: default
execute:
  keep-md: true
---

Research questions:
- What is the effect of a funding agency’s level of selectivity on the uptake of data sharing among research teams which allocate their resources strategically?
- What is the effect of the introduced policy on inequality of resources within the system?


Approach:

1. Reproduce the below analysis from individual data
2. Run across all three types of networks
3. Analyse inequality/how takes up sharing with lagged regression or similar

```{r, echo=FALSE, message=FALSE}
library(patchwork)
library(tidyverse)
library(targets)
library(arrow)

source("R/functions.R")

extrafont::loadfonts(device = "win")
theme_set(theme_bw(base_family = "Hind"))

options(dplyr.summarise.inform = FALSE)

df_all <- open_dataset(tar_read(selectivity_file),
                   format = "csv", schema = schema_funder_selectivity, skip = 1,
                   convert_options = CsvConvertOptions$create(
                     null_values = "<RuntimePrimitiveException>")
                   )

df <- select_funder_selectivity(df_all)

```


# Effect of grant size
```{r fig.width=7, fig.height=9}
#| label: fig-funder-selectivity
#| fig-cap: Gini index and % of groups sharing data dependent on grant size


pdata <- df %>% 
  # remove this value. it was mainly kept for compatibility with prior analyses,
  # but results are very similar to the run with setting -4, so no additional
  # value to show it here
  filter(max_initial_utility != -3, step > 0) %>% 
  group_by(step, funded_share, max_initial_utility) %>% 
  summarise(mean_gini = mean(resources_gini),
            mean_cumulative_gini = mean(total_funding_gini),
            mean_sharing = mean(perc_sharing)) %>% 
  collect() %>% 
  mutate(max_initial_utility = recode(max_initial_utility, `4` = "Uniform initial sharing effort",
                                      `-4` = "Low initial sharing effort"),
         funded_share = scales::percent(funded_share, accuracy = 1))

n_row <- 1

p1 <- pdata %>%  
  ggplot(aes(step, mean_gini, colour = as.factor(funded_share))) +
  geom_line() +
    facet_wrap(vars(max_initial_utility), nrow = n_row) +
    labs(colour = "% of teams receiving funding",
       y = "Gini of current resources", x = NULL)

p2 <- pdata %>%  
  ggplot(aes(step, mean_cumulative_gini, colour = as.factor(funded_share))) +
  geom_line() +
    facet_wrap(vars(max_initial_utility), nrow = n_row) +
    labs(colour = "% of teams receiving funding",
       y = "Gini of total resources",
       x = "Step")

p3 <- pdata %>%  
  ggplot(aes(step, mean_sharing, colour = as.factor(funded_share))) +
  geom_line() +
  facet_wrap(vars(max_initial_utility), nrow = n_row) +
  labs(colour = "% of teams receiving funding",
       y = "% of teams sharing data", x = NULL) 

p3 / p1 / p2 +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "A") & theme(legend.position = "top")
```


