---
title: "Network analysis"
format: 
  html:
    code-fold: true
    fig-dpi: 600
---


```{r, echo=FALSE, message=FALSE}
library(sparklyr)
library(ggplot2)
library(patchwork)
library(dplyr)
library(arrow)
library(here)

theme_custom <- function(base_size = 11) {
  half_line = base_size / 2
  rel_size <- 0.9
  theme_light(base_family = "Hind") %+replace%
            theme(strip.background = element_rect(fill = "grey90", colour = NA),
                  strip.text = element_text(
                           colour = "grey10",
                           size = rel(rel_size),
                           margin = margin(rel_size * half_line, rel_size * half_line, 
                                           rel_size * half_line, rel_size * half_line)
                         ))
}
theme_set(theme_custom())

extrafont::loadfonts()

source(here("R/functions.R"))

Sys.setenv(SPARK_HOME = "/home/tklebel/spark-3.4.0-bin-hadoop3/")
Sys.setenv(HADOOP_HOME = "/home/hadoop/hadoop-3.3.1")
Sys.setenv(HADOOP_CONF_DIR = "/home/hadoop/hadoop-3.3.1/etc/hadoop")
Sys.setenv(YARN_HOME = "/home/hadoop/hadoop-3.3.1")
Sys.setenv(YARN_CONF_DIR = "/home/hadoop/hadoop-3.3.1/etc/hadoop")
Sys.setenv(JAVA_HOME = "/usr/lib/jvm/java-1.11.0-openjdk-amd64")

config <- spark_config()
config$spark.executor.cores <- 5
config$spark.executor.instances <- 20
config$spark.executor.memory <- "20G"
config$spark.hadoop.mapreduce.fileoutputcommitter.algorithm.version <- 2

sc <- spark_connect(master = "yarn", config = config,
                    app_name = "sharing_incentive")

df <- spark_read_csv(
  sc, "vary_incentives",
  path = "/tklebel/data_sharing_abm/vary_incentives.csv.bz2",
  memory = TRUE
)


high_clustering <- spark_read_parquet(
  sc, "high_clustering",
  path = "/tklebel/data_sharing_abm/vary_incentives_individuals_clustered_re_arranged.parquet",
  memory = TRUE
)

```


# Funding incentive part
```{r}
funding_incentive <- df %>% 
  filter(step > 0, fundedshare == .1) %>% 
  group_by(step, sharingincentive, maxinitialutility, network) %>% 
  summarise(n = n(),
            mean_sharing = mean(sharing),
            sd_sharing = sd(sharing)) %>% 
  collect() %>% 
  mutate(maxinitialutility = recode(maxinitialutility, `4` = "Uniform initial sharing effort",
                                      `-4` = "Low initial sharing effort"),
         sharingincentive = scales::percent(sharingincentive, accuracy = 1))
```

```{r}
pdata_funding_incentive <- funding_incentive %>% 
  filter(maxinitialutility == "Low initial sharing effort") %>% 
  pivot_longer(cols = contains("mean"),
               values_to = "y",
               names_to = "outcome_var",
               names_pattern = "mean_(.*)") %>% 
  mutate(
    ymin = case_when(
      outcome_var == "sharing" ~ y - sd_sharing
    ), 
    ymax = case_when(
      outcome_var == "sharing" ~ y + sd_sharing
    )
  ) %>% 
  select(-starts_with("sd"))
```

```{r, fig.width=8, fig.height=5}
#| label: fig-XX

p1 <- pdata_funding_incentive %>% 
  mutate(network = factor(
    network,
    levels = c("none", "random", "clustered", "fragmented"),
    labels = c("No network", "Random network",
               "High clustering", "Low clustering"))) %>% 
  ggplot(aes(step, group = as.factor(sharingincentive))) +
  geom_ribbon(aes(ymin = ymin, ymax = ymax),
              fill = "grey40", alpha = .2, show.legend = FALSE) +
  geom_line(aes(y = y, colour = as.factor(sharingincentive))) +
  facet_wrap(vars(network), nrow = 1) +
  colorspace::scale_colour_discrete_sequential(palette = "Viridis") +
  theme(legend.position = "top") +
  labs(y = "% of teams sharing data", colour = "Sharing incentive", x = NULL) 
```



# Funding selectivity part
```{r}
funding_selectivity <- df %>% 
  filter(step > 0, sharingincentive == .4) %>% 
  group_by(step, fundedshare, maxinitialutility, network) %>% 
  summarise(n = n(),
            mean_sharing = mean(sharing),
            sd_sharing = sd(sharing)) %>% 
  collect() %>% 
  mutate(maxinitialutility = recode(maxinitialutility, `4` = "Uniform initial sharing effort",
                                      `-4` = "Low initial sharing effort"),
         fundedshare = scales::percent(fundedshare, accuracy = 1))
```

```{r}
pdata_funding_selectivity <- funding_selectivity %>% 
  filter(maxinitialutility == "Low initial sharing effort") %>% 
  pivot_longer(cols = contains("mean"),
               values_to = "y",
               names_to = "outcome_var",
               names_pattern = "mean_(.*)") %>% 
  mutate(
    ymin = case_when(
      outcome_var == "sharing" ~ y - sd_sharing
    ), 
    ymax = case_when(
      outcome_var == "sharing" ~ y + sd_sharing
    )
  ) %>% 
  select(-starts_with("sd"))
```

```{r, fig.width=8, fig.height=5}
#| label: fig-YYY

p2 <- pdata_funding_selectivity %>% 
  mutate(network = factor(
    network,
    levels = c("none", "random", "clustered", "fragmented"),
    labels = c("No network", "Random network",
               "High clustering", "Low clustering"))) %>% 
  ggplot(aes(step, group = as.factor(fundedshare))) +
  geom_ribbon(aes(ymin = ymin, ymax = ymax),
              fill = "grey40", alpha = .2, show.legend = TRUE) +
  geom_line(aes(y = y, colour = as.factor(fundedshare))) +
  facet_wrap(vars(network), nrow = 1) +
  colorspace::scale_colour_discrete_sequential(palette = "Inferno") +
  theme(legend.position = "top") +
  labs(y = "% of teams sharing data", colour = "% of teams receiving funding", 
       x = NULL) 
```


```{r, fig.width=10, fig.height=8}
#| label: fig-network-conditions
#| fig-cap: "Effect of network conditions on sharing uptake. All lines display
#| the mean across 100 model runs, with the shaded area as +- 1 standard
#| deviation. (A) Varying levels of sharing incentive assuming funding 
#| selectivity of 0.1. (B) Varying levels of funding selectivity, assuming 
#| sharing incentive of 0.4."
p1 / p2 +
  plot_annotation(tag_levels = "A")
```

# Sub-analysis for 30% incentive with high clustering
Question: why is the rate of sharing here much higher than in lower or higher 
incentives?

## Effort development
```{r}
high_clustering_intervention <- high_clustering %>% 
  filter(fundedshare == .1, maxinitialutility == -4)

high_clustering_effort <- high_clustering_intervention %>% 
  group_by(step, sharingincentive, funded) %>% 
  summarise(mean_effort = mean(effort),
            sd_effort = sd(effort)) %>% 
  collect()
```

```{r}
plot_effort <- function(df) {
  df %>% 
    ggplot(aes(step, mean_effort, group = funded)) +
    geom_ribbon(aes(ymin = mean_effort - sd_effort,
                    ymax = mean_effort + sd_effort), 
                  fill = "grey40", alpha = .15, show.legend = FALSE) +
    geom_line(aes(y = mean_effort, colour = funded)) +
    facet_wrap(vars(sharingincentive), nrow = 2) +
    colorspace::scale_colour_discrete_qualitative() +
    theme(legend.position = "top")
}
```


```{r, fig.width=7, fig.height=6}
#| label: fig-effort-no-network
#| fig-cap: Effort of teams averaged across 100 runs. 
better_pal <- colorspace::sequential_hcl(5, palette = "Inferno")

p2 <- high_clustering_effort %>% 
  mutate(
    funded = factor(
      funded, levels = c(TRUE, FALSE),
      labels = c("Funded teams", "Unfunded teams")
      )) %>% 
  ggplot(aes(step, mean_effort, group = funded)) +
  geom_ribbon(aes(ymin = mean_effort - sd_effort,
                  ymax = mean_effort + sd_effort), 
                fill = "grey40", alpha = .15, show.legend = FALSE) +
  geom_line(aes(y = mean_effort, colour = funded)) +
  facet_wrap(vars(sharingincentive), nrow = 3) +
  coord_cartesian(xlim = c(0, 1500)) +
  scale_colour_manual(values = c(better_pal[2], better_pal[4])) +
  theme(legend.position = c(.85, .15)) +
  labs(y = "Effort of teams", x = NULL, colour = NULL)
```

```{r}
#| label: fig-effort-under-high-clustering
p2
```


## Correlations
```{r}
correlations_high_clustering <- high_clustering_intervention %>% 
  group_by(step, sharingincentive, run_number) %>% 
  summarise(cor_funding = cor(as.numeric(funded), as.numeric(funded_lag)),
            cor_init_resources = cor(as.numeric(funded), as.numeric(initial_resources))) %>% 
  group_by(step, sharingincentive) %>% 
  summarise(mean_cor_funding = mean(cor_funding),
            sd_cor_funding = sd(cor_funding),
            mean_cor_init_resources = mean(cor_init_resources),
            sd_cor_init_resources = sd(cor_init_resources)) %>% 
  collect()
```


```{r}
pdata <- correlations_high_clustering %>% 
  pivot_longer(starts_with("mean"), values_to = "y", names_pattern = "mean_(.*)",
               names_to = "outcome_var") %>% 
  mutate(sharingincentive = scales::percent(sharingincentive, accuracy = 1),
          ymin = case_when(
            outcome_var == "cor_funding" ~ y - sd_cor_funding,
            outcome_var == "cor_init_resources" ~ y - sd_cor_init_resources
          ), 
          ymax = case_when(
            outcome_var == "cor_funding" ~ y + sd_cor_funding,
            outcome_var == "cor_init_resources" ~ y + sd_cor_init_resources
          ),
         outcome_var = case_match(
           outcome_var,
           "cor_funding" ~ "Funding vs. previous funding",
           "cor_init_resources" ~ "Funding vs. initial resources"
          ))
```


```{r, fig.width=10, fig.height=7}
#| label: fig-path-dependency-under-high-clustering
pdata %>% 
  ggplot(aes(step, y, group = sharingincentive)) +
    geom_ribbon(aes(ymin = ymin, ymax = ymax),
                fill = "grey40", alpha = .2, show.legend = FALSE) +
    geom_line(aes(colour = sharingincentive), show.legend = TRUE) +
    facet_wrap(vars(outcome_var), nrow = 2) +
    colorspace::scale_colour_discrete_sequential(palette = "Viridis") +
    labs(x = NULL,
         y = "Correlation (Pearson)",
         colour = "Sharing incentive") +
  theme(legend.position = "top")
```

Main answer: there seems to be a sweetspot between costs and payoff. There is
lower short-term path-dependency. It seems to enable a higher portion of teams
keeping up effort.

Since this is also present for random and low clustering networks, this is 
unlikely to be an artefact of the network structures themselves.
